Graph-Based Neural Reconstruction from Skeletonized 3D Networks

### 2018-03-05 papers/eccv2018/2_method.tex
methods: major updates to skeletonization and neural networks

\begin{figure}[t]
	\centering
	\includegraphics[width=0.32\linewidth]{./figures/skeleton1.png}
	\includegraphics[width=0.32\linewidth]{./figures/skeleton2.png}
	\includegraphics[width=0.32\linewidth]{./figures/skeleton3.png}
	\caption{Example skeletons (in black) extracted from segments using the TEASER algorithm.}
	\label{fig:skeletonization}
\end{figure}

There are two types of errors that can occur in connectomics segmentation.
The first, called a split error, occurs when there are two segments that should have been merged. The second, called a merge error, happens when one segment should be split into two. Generally, it is much more difficult to correct merge errors than to correct split errors,
as the space of possible split proposals grows quickly~\cite{parag2015properties}.
Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. Our method takes as input over-segmentations of EM image volumes generated by state-of-the-art connectomics reconstruction pipelines (Sec.~\ref{sec:neuroproof}). Our goal is to identify locations of split errors and merge the corresponding segments automatically.

From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. The nodes correspond to label segments from the segmentation with edges between segments considered for merging. Ideally, our graph has edges corresponding to all of the segments that were erroneously split. To compute this graph we generate a skeleton for every segment in the pixel-based segmentation (Fig.~\ref{fig:overview}). The skeletonized 3D network is a simplified representation of the overall branching structure of the neurons. From these skeletons we identify potential merge locations and produce the corresponding edges for the graph. To find actual merges we run a classification CNN to generate edge weights corresponding to merge probabilities. We then use a multicut algorithm to generate a partition on the graph where nodes in the same partition are assigned the same output label in the improved segmentation. We will now discuss the three major components to our framework (graph creation, edge weights assignment, and graph partitioning) in more detail.

%\subsection{Graph Creation}

%We generate nodes $N$ and edges $E$ to apply a graph-based optimization strategy for segmentation.
%In addition, these edges receive non-negative weights.

@@

The simplest node generation strategy creates one node for every unique segment label in the input volume. However, some of the millions of labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. It is difficult to extract useful shape features from these segments because of their small, often random, shape. We prune these nodes from the graph by removing all segments with fewer than a threshold $t_{seg} = 20,000$ voxels. This removed on average 56\% of the segments in our  datasets (Sec.~\ref{sec:dataset}). Despite the large number of segments, these regions only take up 1.6\% of the total volume on average.

@@

@@

First, we extract a skeleton from each segment in the label volume using the TEASER algorithm~\cite{sato2000teasar,zhao2014automatic}. Fig.~\ref{fig:skeletonization} shows an example of three extracted skeletons (in black). These skeletons consist of a sequence of \textit{joints}, i.e., locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. We prune the joints that are within $t_{jnt} = 50$ voxels of each other to reduce unnecessary branching. We refer to joints that have only one connected neighbor as \textit{endpoints}. Many of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). We make use of this fact to find merge candidates with the following two-pass pruning algorithm.

In the first pass, we iterate over all endpoints $e$ belonging to a segment $S$ and create a set of segments $\mathbb{S}_e^\prime$ that includes all labels that are within $t_{low}$ voxels from $e$. Elements of $\mathbb{S}_e^\prime$ are candidates for merging. However, this first pass often leads to too many candidates, requiring an additional pass for further pruning. In the second pass, we consider all of the segments in $\mathbb{S}_e^\prime$ for every endpoint $e$. If a segment $S^\prime \in \mathbb{S}_e^\prime$ has an endpoint within $t_{high}$ voxels of $e$, the segment $S$ and $S^\prime$ are considered for merging. We store the midpoints between the two endpoints as the center of the potential merges in the set $\mathbb{S}_c$. This algorithm produces a set of segments to consider for merging. Only these pairs have a corresponding edge in the constructed graph.

@@

\subsection{Edge Weights Assignment}
We assign edge weights $w_e$ to each edge where the weight corresponds to the probability that two nodes belong to the same neuron.
Instead of using handcrafted features to compute the similarity between adjacent nodes, we train a 3D CNN classifier to learn from the manually labeled oversegmentation input volume (Sec.~\ref{sec:dataset}).
If the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}$, where $\beta$ is a tunable parameter that encourages over- or under-segmentation.
%Our  edge weights  correspond to probabilities that two endpoints $e$ in $\mathbb{S}_c$ belong to the same neuron and should be merged. \hp{check}

We extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. The input in all of the channels is in the range $\{-0.5, 0.5\}$. The first channel is $0.5$ only if the corresponding voxel has label $l_1$. The second channel is $0.5$ only if the corresponding voxel has label $l_2$. The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.

@@

%Fig.~\ref{fig:architecture} provides an overview of our CNN architecture.
We use the CNN architecture by Chatfield et al.~\cite{chatfield2014return}. It consists of three layers of double convolutions followed by a max pooling step. The first two max pooling layers are anisotropic with pooling only in the $x$ and $y$ dimensions. The output of this final pooling step is flattened into a 1D vector that is input into two fully connected layers. The final layer produces probabilities with a sigmoid activation function~\cite{funahashi1989approximate}. All of the other activation functions are LeakyReLU~\cite{maas2013rectifier}.

@@

%\begin{figure*}[t]
%	\centering
%	\includegraphics[width=0.85\linewidth]{figures/architecture.png}
%	\caption{The architecture for our classification CNN uses double convolutions followed by max pooling. The number of filters doubles each layer, with a final fully connected layer and sigmoid activation function.}
%	\label{fig:architecture}
%\end{figure*}

@@

After constructing the 3D graph we apply graph partitioning using multicut to compute the final segmentation. 
Using top-down graph partitioning allows us to apply biological constraints on the output. 
Formulating the problem as a multicut minimization problem allows us to guarantee \TODO{FIX} globally consistent solutions where every cycle of ``merge" edges is closed \TODO{CONFUSING}.
To solve this constraining multicut problem we use the method by Keuper et al.~\cite{keuper2015efficient} that produces a feasible solution by greedy additive edge contraction.
We extend the multicut formulation the ``lifted" version of the problem where non-adjacent nodes have edge weights corresponding to the probability of belonging to the same neuron \TODO{(CITE)}.
We determine the lifted edge weights following the example of Keuper et al. \cite{keuper2015efficient}.
\TODO{find better source and describe the process}

>>>>

There are two types of errors that can occur in connectomics segmentations: \textit{split errors} and \textit{merge errors}. 
In a split error, two segments should have been merged into one neuronal process. 
In a merge error, one segment corresponds to more than one neuron.
Generally, it is much more difficult to correct merge errors than to correct split errors,
as the space of possible split proposals grows quickly~\cite{parag2015properties}.
Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. 
Our method takes as input over-segmentations of EM image volumes generated by scalable, state-of-the-art connectomics reconstruction pipelines (Sec.~\ref{sec:neuroproof}). 
Our goal is to identify locations of split errors and merge the corresponding segments automatically.

From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. 
The nodes correspond to labeled segments from the segmentation.
There are edges between all nodes that we consider as merge candidates.
Ideally, our graph has edges corresponding to all of the segments that were erroneously split. 
To compute this graph we generate a skeleton for every segment in the pixel-based segmentation (Fig.~\ref{fig:overview}). 
The skeletonized 3D network is a simplified representation of the overall branching structure of the neurons. 
From these skeletons we identify potential merge locations and produce the corresponding edges for the graph. 
To find actual merges we run a classification CNN to generate edge weights corresponding to merge probabilities. 
We then use a multicut algorithm to generate a partition on the graph where nodes in the same partition are assigned the same output label in the improved segmentation. 
We will now discuss the three major components to our framework (graph creation, edge weight assignment, and graph partitioning) in more detail.

@@

The simplest node generation strategy creates one node for every unique segment label in the input volume. 
However, some of the millions of labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. 
It is difficult to extract useful shape features from these segments because of their small, often random, shape. 
We prune these nodes from the graph by removing all segments with fewer than a threshold $t_{seg} = 20,000$ voxels (\TODO{~XX\SI{}{\nano\meter^3}}). 
This removed on average \TODO{XX}\% of the segments in our  datasets (Sec.~\ref{sec:dataset}). Despite the large number of segments, these regions only take up \TODO{XX}\% of the total volume on average.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.32\linewidth]{./figures/skeleton1.png}
	\includegraphics[width=0.32\linewidth]{./figures/skeleton2.png}
	\includegraphics[width=0.32\linewidth]{./figures/skeleton3.png}
	\caption{Example skeletons (in black) extracted from segments using a variant of the TEASER algorithm.}
	\label{fig:skeletonization}
\end{figure}

@@

First, we extract a skeleton from each segment in the label volume using a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
This skeletonization algorithm repeatedly uses Dijkstra's algorithm to find the farthest voxel from a seed location. 
Since this algorithm is non-linear in the number of voxels, we downsample the datasets so that there are voxel samples every $\SI{30}{\nano\meter}$ in each dimension.
Empirically, this reduced the running time for skeletonization by $\TODO{XX\%}$, with minimal reduction in skeleton accuracy ($\TODO{XX\%}$ fewer branches). 
Fig.~\ref{fig:skeletonization} shows three examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, i.e., locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We prune the joints that are within $t_{jnt} = \TODO{XX}\SI{}{\nano\meter}$ of each other to reduce unnecessary branching. 
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
Many of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). 
We make use of this fact to find merge candidates with the following two-pass pruning algorithm.

In the first pass, we iterate over all endpoints $e$ belonging to a segment $S$ and create a set of segments $\mathbb{S}_e^\prime$ that includes all labels that are within $t_{low}$ voxels from $e$.
Elements of $\mathbb{S}_e^\prime$ are candidates for merging. 
However, this first pass often leads to too many candidates, requiring an additional pass for further pruning. 
In the second pass, we consider all of the segments in $\mathbb{S}_e^\prime$ for every endpoint $e$. 
If a segment $S^\prime \in \mathbb{S}_e^\prime$ has an endpoint within $t_{high}$ voxels of $e$, the segment $S$ and $S^\prime$ are considered for merging. 
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. Only these pairs have a corresponding edge in the constructed graph.

\subsection{Edge Weights Assignment}
We assign edge weights $w_e$ to each edge where the weight corresponds to the probability that two nodes belong to the same neuron.
We train a 3D CNN classifier to learn from the manually labeled oversegmentation input volume (Sec.~\ref{sec:dataset}).
If the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}$, where $\beta$ is a tunable parameter that encourages over- or under-segmentation.

We extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. 
The input in all of the channels is in the range $\{-0.5, 0.5\}$. 
The first channel is $0.5$ only if the corresponding voxel has label $l_1$. 
The second channel is $0.5$ only if the corresponding voxel has label $l_2$. 
The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This significantly reduces the need for generating costly manually-labeled ground truth.

@@

@@

We use \textit{VGG-style} blocks as presented by Chatfield et al.~\cite{chatfield2014return}. 
Each blocks consists of two double convolutions with filter size $3\times3\times3$ followed by a max pooling step. 
There are three layers of these blocks. 
The first two max pooling layers are anisotropic with pooling only in the $x$ and $y$ dimensions. 
The output of this final pooling step is flattened into a 1D vector that is input into two fully connected layers. 
The final layer produces probabilities with a sigmoid activation function~\cite{funahashi1989approximate}. 
All of the other activation functions are LeakyReLU~\cite{maas2013rectifier}.

@@

@@

After constructing the 3D graph we apply graph partitioning with a multicut formulation to compute the final segmentation. 
Using this global graph partitioning allows us to apply biological constraints on the output. 
The multicut minimization problem allows us to guarantee globally consistent solutions where every cycle is ``closed'' under merging.
To solve this constraining multicut problem we use the method by Keuper et al.~\cite{keuper2015efficient} that produces a feasible solution by greedy additive edge contraction.
We extend the multicut formulation to the ``lifted'' version of the problem where non-adjacent nodes have ``lifted'' edges.
These edges penalize or reward distance nodes for belonging to the same segment and produce benefits over traditional multicut. 
We determine the lifted edge weights following the example of Beier et al~\cite{beier2017multicut}.
We run the traditional multicut algorithm on our graph varying a bias parameter that encourages over- and under-segmentation. 
\TODO{ADD ADDITIONAL FEATURES}
